---
title: "3D Plots"
author: "Anastasia Sorokina"
output:
  html_document:
    df_print: paged
---

# PLOTLY package

```{r}
library(plotly)
```
## Basic 3D Surface Plot

```{r}
fig <- plot_ly(z = ~volcano)
fig <- fig %>% add_surface()
fig
```

## 2D Kernel Density Estimation

```{r}
kd <- with(MASS::geyser, MASS::kde2d(duration, waiting, n = 50))
fig <- plot_ly(x = kd$x, y = kd$y, z = kd$z) %>% add_surface()
fig
```

# RGL package

```{r}
library(rgl)
```

```{r}
data(iris)
head(iris)
```

```{r}
x <- sep.l <- iris$Sepal.Length
y <- pet.l <- iris$Petal.Length
z <- sep.w <- iris$Sepal.Width
```

```{r}
rgl.open()
```

## Scatter plots

```{r}
rgl.points(x, y, z, color ="lightgray")
```

```{r}
rgl.open()
rgl.bg(color = "white") 
rgl.points(x, y, z, color = "pink", size = 5) 
```

### Draw spheres

```{r}
rgl.open()
rgl.bg(color = "white") 
rgl.spheres(x, y, z, r = 0.2, color = "grey") 
```

### Add boxes

```{r}
rgl.open()
rgl.bg(color = "white") 
rgl.spheres(x, y, z, r = 0.2, color = "white")  
rgl.bbox(color=c("#333377","black"), emission="#333377",
         specular="#3333FF", shininess=5, alpha=0.8 ) 
```

# CAR package

```{r}
library(car)
```

## 3D plot with the regression plane

```{r}
scatter3d(x = sep.l, y = pet.l, z = sep.w)
```

### Points by groups
```{r}
scatter3d(x = sep.l, y = pet.l, z = sep.w, groups = iris$Species)
```

### Add concentration ellipsoids
```{r}
scatter3d(x = sep.l, y = pet.l, z = sep.w, groups = iris$Species,
          surface=FALSE, grid = FALSE, ellipsoid = TRUE)
```

### Add text labels for the points
```{r}
scatter3d(x = sep.l, y = pet.l, z = sep.w, 
          surface=FALSE, labels = rownames(iris), id.n=nrow(iris))
```

### Save images
```{r}
rgl.snapshot(filename = "plot.png")
rgl.postscript("plot.pdf",fmt="pdf")
```

# plot3D package

```{r}
library(plot3D)
```

## Scatter plots
```{r}
scatter3D(x, y, z, bty = "g", colkey = FALSE, main ="bty= 'g'")
```
```{r}
scatter3D(x, y, z, pch = 18, bty = "u", colkey = FALSE, 
   main ="bty= 'u'", col.panel ="steelblue", expand =0.4, 
   col.grid = "darkblue")
```

## 3D Histogram

```{r}
data(VADeaths)
#  hist3D and ribbon3D with greyish background, rotated, rescaled,...
hist3D(z = VADeaths, scale = FALSE, expand = 0.01, bty = "g", phi = 20,
        col = "thistle", border = "white", shade = 0.2, ltheta = 90,
        space = 0.3, ticktype = "detailed", d = 2)
```

### Add interactivity

```{r}
library("plot3Drgl")
plotrgl()
```

# Explore other options
```{r}
library(rayshader)
library(ggplot2)
library(tidyverse)
```
## 3D Surface Plots

```{r}
rgl.open()
gg = ggplot(diamonds, aes(x, depth)) +
  stat_density_2d(aes(fill = stat(nlevel)), 
                  geom = "polygon",
                  n = 100,bins = 10,contour = TRUE) +
  facet_wrap(clarity~.) +
  scale_fill_viridis_c(option = "A")
plot_gg(gg,multicore=TRUE,width=5,height=5,scale=250)
```

## 3D maps

```{r}
library(sf)
library(viridis)
rgl.open()
nc = st_read(system.file("shape/nc.shp", package="sf"), quiet = TRUE)
gg_nc = ggplot(nc) +
  geom_sf(aes(fill = AREA)) +
  scale_fill_viridis("Area") +
  ggtitle("Area of counties in North Carolina") +
  theme_bw()

plot_gg(gg_nc, multicore = TRUE, width = 6 ,height=2.7, fov = 70)
render_depth(focallength=100,focus=0.72)
```


### And many other options

For more details: https://www.tylermw.com/3d-ggplots-with-rayshader/

```{r}
rgl.open()
a = data.frame(x=rnorm(20000, 10, 1.9), y=rnorm(20000, 10, 1.2) )
b = data.frame(x=rnorm(20000, 14.5, 1.9), y=rnorm(20000, 14.5, 1.9) )
c = data.frame(x=rnorm(20000, 9.5, 1.9), y=rnorm(20000, 15.5, 1.9) )
data = rbind(a,b,c)

#Lines
pp = ggplot(data, aes(x=x, y=y)) +
  geom_hex(bins = 20, size = 0.5, color = "black") +
  scale_fill_viridis_c(option = "C")
plot_gg(pp, width = 4, height = 4, scale = 300, multicore = TRUE)

#No lines
pp_nolines = ggplot(data, aes(x=x, y=y)) +
  geom_hex(bins = 20, size = 0) +
  scale_fill_viridis_c(option = "C")
plot_gg(pp_nolines, width = 4, height = 4, scale = 300, multicore = TRUE)
```


```{r}
rgl.open()
mtcars_gg = ggplot(mtcars) + 
  geom_point(aes(x=mpg,color=cyl,y=disp),size=2) +
  scale_color_continuous(limits=c(0,8)) +
  ggtitle("mtcars: Displacement vs mpg vs # of cylinders") +
  theme(title = element_text(size=8),
        text = element_text(size=12)) 

plot_gg(mtcars_gg, height=3, width=3.5, multicore=TRUE, pointcontract = 0.7, soliddepth=-200)
```


```{r}
library(tidyverse)
rgl.open()
measles = read_csv("https://tylermw.com/data/measles_country_2011_2019.csv")
melt_measles = reshape2::melt(measles, id.vars = c("Year", "Country", "Region", "ISO3"))
melt_measles$Month = melt_measles$variable
melt_measles$cases = melt_measles$value
melt_measles %>% 
  group_by(Year, Month) %>%
  summarize(totalcases = sum(cases,na.rm = TRUE)) %>% 
  mutate(totalcases = ifelse(Year == 2019 & !(Month %in% c("January","February","March")), NA, totalcases)) %>%
  ggplot() + 
  geom_tile(aes(x=Year, y=Month, fill=totalcases,color=totalcases),size=1,color="black") + 
  scale_x_continuous("Year", expand=c(0,0), breaks = seq(2011,2019,1)) +
  scale_y_discrete("Month", expand=c(0,0)) +
  scale_fill_viridis("Total\nCases") +
  ggtitle("Reported Worldwide Measles Cases") +
  labs(caption = "Data Source: WHO") +
  theme(axis.text = element_text(size = 12),
        title = element_text(size = 12,face="bold"),
        panel.border= element_rect(size=2,color="black",fill=NA)) -> 
measles_gg

plot_gg(measles_gg, multicore = TRUE, width = 6, height = 5.5, scale = 300,shadowcolor = "#3a4f70")
```

